\documentclass[letterpaper, 10pt]{article}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,scrextend}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{silence}
\WarningFilter{latex}{You have requested package}
\input{ltx/pkg/preamble}





\begin{document}

\lhead{MAT224 Linear Algebra II}
\chead{Review Session}
\rhead{Week 12}

\title{Linear Algebra II \\ \Large{MAT224}}
\author{Lennart Döppenschmitt}
% \maketitle
% \tableofcontents

\section*{Review Session}%
\textbf{Textbook:} All of it! \footnote{Everything we covered.}


\lb
\textbf{Discussion 1}
\lb
\begin{enumerate}
    \item Let $U \subseteq \pol{n}{\bb C}$ be the subset of polynomials that
        satisfy $p'(x) =x \cdot p(x)$. Explain why this is a subspace or not.
    \item Let $W$ be the line of points in $ \ttpl{x}{y} ∈ \bb C^2$ that satisfy the equation
        \[ 5x - 3iy + 2 = 0 \]
        Explain why this is a subspace or not.
\end{enumerate}

\newpage
\lb
\textbf{Discussion 2}
\lb
\begin{enumerate}
    \item Prove that $U$ is a subspace of $V$ if and only if $\spn{U} = U$.
    \item We have seen that $\spn{S} = \spn{R}$ for families of vectors $S$ and $R$ in a
        vector space $V$ if $S \subseteq \spn{R}$ and $R \subseteq \spn{S}$.
        Is the converse true as well?
    \item Give an example where $R \neq S$, but $\spn{R} = \spn{S}$.
\end{enumerate}


\newpage
\lb
\textbf{Discussion 3}
\lb
Why can the vector space $\mat{2}{\bb C}$ of $2 \times 2$ matrices be written as a direct sum
of symmetric and antisymmetric matrices?
\[ \mat{2}{\bb C} = \mathcal{S} \oplus \mathcal{A} \]


\newpage
\lb
\textbf{Discussion 4}
\lb
Suppose $V = U \oplus W$ and $S$ and $R$ are linearly independent families of vectors in $U$
and $W$ respectively. Show that $S \cup R$ is linearly independent in $V$.

\newpage
\lb
\textbf{Discussion 5}
\lb
Complete the family of polynomials $ \cb{1 - x ~,~ 1 - x + x^2} $
to find a basis of $\pol{3}{\R}$.

\newpage
\lb
\textbf{Discussion 6}
\lb
\begin{enumerate}
    \item  Let $T$ be a transformation from $V$ to $W$. Is the preimage
        $\iv T (\vec y)$ for $\vec y ∈ W$ a subspace of $V$?
    \item Fix one vector $\vec x_0 ∈ \iv T (\vec y)$. Show that any other vector
        $\vec x ∈ \iv T (\vec y)$ can be written as
        \[ \vec x = \vec x_0 + \vec k \]
        for some $\vec k ∈ \tx{ker}(T)$.
\end{enumerate}

\newpage
\lb
\textbf{Discussion 7}
\lb
Let $T$ be a linear transformation on $V$.
\pr Explain why the dimension of the
kernel $\ker(T)$ is equal to $\dim{V} - \tx{rank}([T]_α^α)$.


\newpage
\lb
\textbf{Discussion 8}
\lb
All matrices $A$ in $\mat{2}{\bb C}$ are similar to either
$ \begin{pmatrix} λ_1 & 0 \\ 0 & λ_2 \end{pmatrix} $ or
$ \begin{pmatrix} λ & 1 \\ 0 & λ \end{pmatrix} $.




\newpage
\lb
\textbf{Discussion 9}
\lb
If $N$ is a nilpotent transformation on $V$, then $\ker(N^j) < \ker(N^{j+1})$
for all $j < \tx{index}(N)$.

\newpage
\lb
\textbf{Discussion 10}
\lb
Compute the inverse of the left-shift transformation on a vector space $V$ in a given a basis
$α$.


\newpage
\lb
\textbf{Discussion 11}
\lb
At how many points do we need to evaluate a polynomial of degree $3$  to reconstruct it?
Explain your answer in terms of a linear transformation
\begin{align*}
    \pol{3}{\R} &\ra \R^n \\
    p(x) &\mapsto \tttpl{p(x_1)}{\vdots}{p(x_n)}
\end{align*}


\newpage
\lb
\textbf{Discussion 12}
\lb
Two eigenvectors $\vec v$ and $\vec w$ of $T$ with different eigenvalues $λ$ and $σ$ are
linearly independent.

\newpage
\lb
\textbf{Discussion 13}
\lb
What are the real and imaginary parts of $\frac{1}{a + ib}$?



\newpage
\lb
\textbf{Discussion 14}
\lb
Consider $V = \R$ as a vector space over $\Q$.
\begin{enumerate}
    \item Is the family of vector $ \cb{1, \sqrt{2}}$ linearly independent or dependent?
    \item What is the dimension of the subspace $\spn{1, \sqrt{2}, \sqrt{3}, \sqrt{4}}$?
    \item What do the above results suggest about $\dim(V)$?
\end{enumerate}

\end{document}


